---
title: "Cleaning to Pare Down File"
author: "Katrina Gutierrez"
date: "10/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cleaning Data

Goal: Clean data everything_merged.csv to produce a smaller file with:
- dates for the weeks
- limited predictor and outcome variables that are available weeks 1-33
    - exception: SCHLHRS - weeks 13-33


The completed file has numbers for the responses (For example, 1-4 instead of the "blah")



```{r }
library("readr")
library("data.table")
library("tidyverse")

```

Loading the data

```{r }

# need to do fread instead of read.csv for the large file 
df <- fread("everything_merged.csv")
dates <- fread("dates.csv")


```

Cleaning steps:
- Select variables that are present weeks 1-33 (+ SCHLHRS which is weeks 13-33)
- Add date variable
- Exporting csv


```{r }

predictors <- c("EEDUC","RHISPANIC", "RRACE", "THHLD_NUMKID", "THHLD_NUMPER", "INCOME", "CURFOODSUF", "TSPNDFOOD", "TSPNDPRPD", "MORTCONF", "EST_ST")
outcomes <- c("TEACH1", "TEACH2", "TEACH3", "TEACH4", "TEACH5", "COMPAVAIL", "INTRNTAVAIL", "SCHLHRS")


clean <- df %>% 
  dplyr::select(WEEK, PWEIGHT, predictors, outcomes) %>%
  filter(WEEK <= 33)

rm(df)

clean <- merge(dates, clean, by = "WEEK")  

write.csv(clean,"clean_data.csv", row.names = FALSE)
```


